---
title: rlstats Package
---

The fourth project in CSMC405 involved creating an R package. The focus of this package was open ended, and could be geared towards anything we wanted to choose. I decided to create an R package to analyze gameplay data from the video game Rocket League. Find below some examples of the three functions contained in my package. For full details on my package, check out the [package website](https://adc-405-s26.github.io/rlstats/). You can also find the source code for this package [here](https://github.com/ADC-405-S26/rlstats).

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rlstats

<!-- badges: start -->
<!-- badges: end -->

## Installation

You can install the development version of rlstats from
[GitHub](https://github.com/) with:

``` r
#install.packages("remotes")
#install.packages("devtools")
devtools::install_github("ADC-405-S26/rlstats")
```

## Examples

This package contains three main functions. Two for analyzing a specific player, and one for comparison between teams in a match. Below you'll see usage of all three.

```{r example}
library(rlstats)
```


#### stats_for_player

This function outputs a table with some basic summary statistics for a specified player

```{r}
stats_for_player(TDMZ443games, name = "TDMZ443")
```

If your dataset contains data from several games, the above function call will return aggregated statistics for all those games. If you want to specify a particular game, you pass in its match guid.

```{r}
stats_for_player(TDMZ443games, name = "TDMZ443", game = '8DF7C4C211F1560E0ED84EA271D83282')
```


#### speed_boost_plots

This function returns two plots displaying boost total and speed information for a specified player throughout the course of a specified game

```{r}
speed_boost_plots(TDMZ443games, name = 'TDMZ443', game = '8DF7C4C211F1560E0ED84EA271D83282')
```


#### compare_teams

This functions outputs a table with some summary statistics allowing comparison between the performance of both teams in specified match

```{r}
compare_teams(TDMZ443games, '8DF7C4C211F1560E0ED84EA271D83282')
```
