---
title: "Mountaineering Data Analysis with SQL"
---

The second project in CMSC405 involved making use of SQL+R to design and create a relational database with at least three different datasets that could be joined with a common key, and then perform data analysis on our database. My partner Laura Henze and I chose to investigate Himalayan Mountaineering data. Find our database creation and exploration below.


```{r setup}
#| message: false
#| warning: false
#| include: false
library(tidyverse)
library(mdsr)
library(DBI)
library(RMariaDB)
library(dbplyr)
library(duckdb)

```


# Accessing data

We will be considering the Himalayan Climbing Expedition datasets available on GitHub.
[Himalayan Expedition Datasets](https://github.com/rfordatascience/tidytuesday/tree/main/data/2020/2020-09-22)

# Create the connection
```{r}

con_exp <- DBI::dbConnect(duckdb::duckdb(), dbdir = "exp_db")

```

# Read in the raw data from Github
```{r}

members <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-09-22/members.csv')
expeditions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-09-22/expeditions.csv')
peaks <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-09-22/peaks.csv')


```

```{r}
glimpse(peaks)

# Quick check to see what the data look like
```

```{r}
summary(peaks)

# Nothing seems like it needs to be removed.
```

```{r}
glimpse(expeditions)

# Quick check to see what the data look like
```

```{r}
summary(expeditions)

#Don't see any issue in keeping everything for now, besides peak_name 
# which will be redundant after join.
```

```{r}
glimpse(members)

# Quick check to see what the data look like
```

```{r}
summary(members)

# Remove some heavily NA columns, and columns that can be gotten from a join 
# (year, etc)
```

```{r}
# Remove some uninteresting/heavily null columns in members.
# Remove also columns that we will be able to pull from the peaks or expeditions
# table with a join to reduce redundancy when joining (peak_name, peak_id, year,
# and season can all be obtained from joining with expeditions and peak)

# Also remove duplicate key which won't play nice with SQL

members_clean <- members %>% select(!c('highpoint_metres','death_cause','death_height_metres','injury_type','injury_height_metres','peak_name','peak_id','year','season')) %>% 
  filter(member_id!='KANG10101-01')
  

glimpse(members_clean) #looks good
```

```{r}
# Must remove related datapoints from expeditions that we did in above chunk
# for members. This is some weird duplicate point that messes things up.
# also remove "peak_name" column. This can be obtained by joining with peaks
# table, and will be redundant when we do that join

expeditions_clean <- expeditions %>% 
  select(-peak_name) %>% 
  filter(expedition_id != 'KANG10101')
```

```{r}
# Write CSV's to read from for database

write_csv(peaks,'peaks.csv')
write_csv(expeditions_clean, 'expeditions.csv')
write_csv(members_clean,'members.csv')
```

### ---------------------------------------------------------------------------

# Designing the database

```{sql connection=con_exp}
/* making sure to drop the tables in case previous versions exist already */
DROP TABLE IF EXISTS members;
DROP TABLE IF EXISTS expeditions;
DROP TABLE IF EXISTS peaks;
```


# Create a table from the members dataset
```{sql connection= con_exp}

CREATE TABLE members (
  expedition_id VARCHAR(20) NOT NULL,
  member_id VARCHAR(30) NOT NULL,
  sex CHAR(1),
  age SMALLINT,
  citizenship VARCHAR(100),
  expedition_role VARCHAR(100),
  hired BOOLEAN,
  success BOOLEAN,
  solo BOOLEAN,
  oxygen_used BOOLEAN,
  died BOOLEAN,
  injured BOOLEAN,

  PRIMARY KEY (member_id)
);

```

# Create a table from the expeditions dataset
```{sql connection= con_exp}

CREATE TABLE expeditions (
  expedition_id VARCHAR(20) NOT NULL,
  peak_id VARCHAR(10) NOT NULL,
  year SMALLINT,
  season VARCHAR(20),

  basecamp_date DATE,
  highpoint_date DATE,
  termination_date DATE,

  termination_reason VARCHAR(255),
  highpoint_metres INT,

  members INT,
  member_deaths INT,
  hired_staff INT,
  hired_staff_deaths INT,

  oxygen_used BOOLEAN,
  trekking_agency VARCHAR(100),

  PRIMARY KEY (expedition_id)
);

```

# Create a table from the peaks dataset
```{sql connection= con_exp}

CREATE TABLE peaks (
  peak_id VARCHAR(10) NOT NULL,
  peak_name VARCHAR(100),
  peak_alternative_name VARCHAR(150),
  height_metres INT,
  climbing_status VARCHAR(50),
  first_ascent_year SMALLINT,
  first_ascent_country VARCHAR(150),
  first_ascent_expedition_id VARCHAR(20),

  PRIMARY KEY (peak_id)
);

```

# Copy each table into the database
```{sql connection= con_exp}

COPY members FROM 'members.csv' (FORMAT CSV, HEADER, NULL 'NA');

```

```{sql connection=con_exp}

COPY expeditions FROM 'expeditions.csv' (FORMAT CSV, HEADER, NULL 'NA');

```

```{sql connection=con_exp}

COPY peaks FROM 'peaks.csv' (FORMAT CSV, HEADER, NULL 'NA');

```

# Check database
```{sql connection=con_exp}
SHOW DATABASES;
```

# Check table variables
```{sql connection=con_exp}
DESCRIBE peaks;
```

```{sql connection=con_exp}
DESCRIBE expeditions;
```

```{sql connection=con_exp}
DESCRIBE members;
```

### ---------------------------------------------------------------------------

# Querying data using SQL

# Some Interesting Tidbits

Average Peak Height by Climbing Status: Compares the average height of peaks across different climbing status categories. 
```{sql connection=con_exp}

/* Calculate the average height of peaks within each climbing
   status category. This allows us to investigate whether
   unclimbed or restricted peaks tend to be taller or shorter
   than peaks that have already been successfully climbed. */

SELECT climbing_status, AVG(height_metres) AS avg_height
FROM peaks
GROUP BY climbing_status
ORDER BY avg_height DESC;
```

We can see that helps taller peaks are actually more likely to be climbed. This suggests that elevation may not be a key factor in difficulty. Some peaks may remain unclimbed due to factors other than height, such as accessibility or restrictions.


Success Over Time: Examines how the probability of a successful climb changes across years.

```{sql connection=con_exp}

/* Calculate the proportion of climbers who successfully summited
   each year. The SUM() counts successful climbers while COUNT(*)
   gives the total number of climbers. Multiplying by 1.0 ensures
   decimal division rather than integer division. */
   
SELECT 
  e.year,
  COUNT(*) AS total_members,
  SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) AS successful_members,
  ROUND(
    SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) * 1.0 / COUNT(*), 
    3
  ) AS success_rate
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
GROUP BY e.year
ORDER BY e.year;
```

We can see that the success rate has a generally upward trend over time with a larger portion of climbers summiting each year. This suggests that this may be due to improvements in gear, technology or training. 


Success VS Oxygen Used: Compares the success rates of those who used supplemental oxygen on their climb vs those who did not. 

```{sql connection=con_exp}

/* Compare success rates between climbers who used supplemental
   oxygen and those who did not. Success rate is computed as the
   number of successful climbers divided by the total climbers in
   each oxygen-use group. */
   
SELECT 
  m.oxygen_used,
  COUNT(*) AS total_climbers,
  SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) AS successes,
  ROUND(
    SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) * 1.0 / COUNT(*), 
    3
  ) AS success_rate
FROM members m
GROUP BY m.oxygen_used
ORDER BY m.oxygen_used;

```

This output tells us that climbers who used supplemental oxygen (oxygen_used = TRUE) had a much higher success rate that those who didn't. This also tells us that many more climbs have been done without the use of oxygen as the number of total climbers that didn't use oxygen is substantially higher. This could be due to limits in oxygen accessibility over time. 


JOIN - Top Countries Participating: Displays which countries are most involved in Himalayan expeditions. 

```{sql connection=con_exp}
/* Join the members and expeditions tables to identify which
   countries contribute the largest number of climbers to
   Himalayan expeditions. We also count the number of distinct
   expeditions each country has participated in to distinguish
   between countries with many climbers on a few expeditions
   and those with broad participation across many expeditions. */
   
SELECT 
  m.citizenship,
  COUNT(*) AS num_climbers,
  COUNT(DISTINCT m.expedition_id) AS expeditions_participated
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
GROUP BY m.citizenship
ORDER BY num_climbers DESC
LIMIT 10;

```
This analysis shows how participation in Himalayan expeditions varies by country, highlighting which nations contribute the most climbers and how widely they are involved across expeditions. Nepal appears as the largest participator, which makes sense given the location of the Himalayas. Also, climbers from Nepal are also often guides which will also contribute to this number. 

# An Analysis of Impact of Hired Staff in Expeditions

Get an idea of the spread of hired staff vs recreational climbers in the expeditions

```{sql connection=con_exp}
   /* Compute summary statistics for both expedition members and
   hired staff to understand the overall distribution of group
   sizes. Examining the minimum, maximum, and average values
   helps determine whether there is substantial variation across
   expeditions and motivates further analysis of member-to-staff
   ratios. */
   
SELECT 
  MAX(members) AS max_members, MIN(members) AS min_members, AVG(members) AS average_members,
  MAX(hired_staff) AS max_staff, MIN(hired_staff) AS min_staff, AVG(hired_staff) AS average_staff
  FROM expeditions

```

We see a pretty big spread, so we'll probably have to dive deeper. Lets ignore those that have zero members (as this would be staff only). 

Check what peaks have the highest average member to staff ratio for expeditions. We will have to exclude expeditions with 0 members or 0 staff to get
a meaningful result. We will check a count how many of these there are too to get
an idea of how much we're missing

```{sql connection=con_exp}

/* Calculate the average climber-to-staff ratio for each peak
   across all recorded expeditions. This ratio provides insight
   into the level of support available to climbers, with larger
   values indicating fewer hired staff relative to the number of
   recreational climbers. Expeditions with zero members or zero
   staff are excluded to avoid undefined ratios and ensure that
   comparisons are meaningful. */
   
SELECT p.peak_name, AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
FROM expeditions e
JOIN peaks p
  ON e.peak_id = p.peak_id
WHERE members != 0 AND hired_staff != 0
GROUP BY p.peak_name
ORDER BY member_staff_ratio DESC

```
We have around 10,000 rows, so the above query only acts on about 70% of our data. I still think this is enough to produce some interesting results
```{sql connection=con_exp}

/* Count the number of expeditions that have either zero members
   or zero hired staff. These records were excluded from the
   member-to-staff ratio analysis because the ratio would be
   undefined or not meaningful. This count helps quantify how
   much of the dataset is being removed and assess whether the
   exclusion could impact the results. */
   
SELECT count(*) as count
FROM expeditions 
WHERE members = 0 OR hired_staff = 0

```
Let's look a bit closer now at success rates. We will exclude the same rows we excluded above to stay consistent

```{sql connection=con_exp}

/* Calculate the expedition success rate for each peak by dividing
   the number of successful expeditions by the total number of
   expeditions recorded for that peak. An expedition is considered
   successful if its termination reason contains the word "Success".
   Expeditions with zero members or zero hired staff are excluded
   to remain consistent with the earlier member-to-staff ratio
   analysis and allow for meaningful comparisons between peaks. */
   
  SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name
  ORDER BY success_rate DESC
```
Seems like a lot of them are 1 or 0. This could mean there are only a few (or one) expeditions.
We will still use this for plotting, it might be interested to see how the member/staff ratio looks for some of these, but for interest sake lets filter these 1's and 0's out to see what the middle looks like.

```{sql connection=con_exp}

/* Filter the peak-level success rates to exclude peaks with
   perfect (100%) or zero success rates. Many of these extreme
   values may result from peaks with only a small number of
   recorded expeditions, making them less informative for
   identifying trends. Focusing on peaks with intermediate
   success rates provides a clearer picture of how expedition
   outcomes vary across more frequently climbed peaks. */

SELECT *
FROM (
  SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name
) t
WHERE success_rate NOT IN (0, 1)
ORDER BY success_rate DESC;


```



### ---------------------------------------------------------------------------

# Some Interesting Visualizations


Death Rate of Himalayan Climbers Over Time: Shows how the death rate of Himalayan climbers has changed over the years. 

```{sql connection=con_exp}

/* Compute annual death rates by treating TRUE deaths as 1 and
   FALSE deaths as 0. The average therefore represents the
   proportion of climbers who died in a given year. We exclude
   years with fewer than 50 climbers to reduce noise from very
   small sample sizes. */

SELECT 
  e.year,
  AVG(CASE WHEN m.died = TRUE THEN 1 ELSE 0 END) AS death_rate,
  COUNT(*) AS total_climbers
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
WHERE e.year IS NOT NULL
GROUP BY e.year
HAVING COUNT(*) > 50
ORDER BY e.year;
```

```{r}
# Execute the SQL query and store the results in an R data frame.
# This allows us to move from database analysis to visualization
# and further processing in R.

death_trend_data <- dbGetQuery(con_exp, "
SELECT 
  e.year,
  AVG(CASE WHEN m.died = TRUE THEN 1 ELSE 0 END) AS death_rate,
  COUNT(*) AS total_climbers
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
WHERE e.year IS NOT NULL
GROUP BY e.year
HAVING COUNT(*) > 50
ORDER BY e.year
")

```

```{r}
# Visualize how annual death rates have changed over time.
# The smooth trend line helps identify long-term patterns
# that may not be obvious from year-to-year variation.

ggplot(death_trend_data, aes(x = year, y = death_rate)) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Death Rate of Himalayan Climbers Over Time",
    x = "Year",
    y = "Death Rate"
  ) +
  theme_minimal()

```
By plotting the proportion of climbers who died each year, we can see that the death rate has steadily declined since 1970, which makes sense as there has been upgrades to technology as well as more guides present on these expeditions. The data increases steadily at the start which makes sense given the smaller number of climbers in the beginning years as well as the lack of knowledge of the climb. 



# Continuation of Hired Staff Impact Analysis

My interest here is to see if there is any positive correlation between these two. 
I expect that there will be some, but not extremely strong, as what really matters
is the difficulty of the peak, the experience of the people, etc. But presumably more professional staff members should result in more success.

```{sql connection=con_exp}

/*Query to get both the success rate and the member-staff ratio for all peaks 
without 100% or 0% success rate*/

 SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate,
    AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name

```

```{r}

# Execute the SQL query and store the results in an R data frame.
# This allows us to move from database analysis to visualization
# and further processing in R.

success_rate_member_ratio_data <- dbGetQuery(con_exp, "
SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate,
    AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name")

```

```{r}

# Visualize the relationship between staff support and
# expedition success. If peaks with lower climber-to-staff
# ratios tend to have higher success rates, we would expect
# to see a positive trend in the data.

ggplot(success_rate_member_ratio_data, aes(x = member_staff_ratio, y = success_rate)) +
  geom_point() +
  geom_smooth(se = FALSE)+
labs(
    title = "Success Rate of Expeditions vs Recreational Climber/Staff Ratio",
    x = "Recreational Climber/Staff Ratio",
    y = "Success Rate"
  ) +
  theme_minimal()
```

```{r}

# Transfer the filtered SQL results into R for visualization.
# Peaks with success rates of exactly 0% or 100% are excluded
# so the analysis can focus on peaks with more variation in outcomes.

success_rate_member_ratio_data_filtered <- dbGetQuery(con_exp, "
 SELECT * FROM
 (SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate,
    AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name
 ) t
WHERE success_rate NOT IN (0, 1)
 ")

```

```{r}

# Visualize the relationship between climber-to-staff ratio and
# success rate after removing peaks with 0% and 100% success rates.
# This helps reveal trends that may be obscured by extreme values.

ggplot(success_rate_member_ratio_data_filtered, aes(x = member_staff_ratio, y = success_rate)) +
  geom_point() +
  geom_smooth(se = FALSE)+
labs(
    title = "Success Rate of Expeditions vs Recreational Climber/Staff Ratio",
    x = "Recreational Climber/Staff Ratio",
    y = "Success Rate"
  ) +
  theme_minimal()
```

Unfiltered version seems to suggest there may be a small correlation. We do see more 100% success rates with higher ratios. However when we glance at the filtered version, it seems like there is less of one (though there may be something small). But overall I imagine other factors like experience, and peak difficulty matter the most.

### ---------------------------------------------------------------------------

```{r}
# terminate the SQL connection

DBI::dbDisconnect(con_exp, shutdown = TRUE)

```

### ---------------------------------------------------------------------------
