---
title: "Project 2"
---

```{r setup}
#| message: false
#| warning: false
#| include: false
library(tidyverse)
library(mdsr)
library(DBI)
library(RMariaDB)
library(dbplyr)
library(duckdb)

```

### ---------------------------------------------------------------------------

# Accessing data

We will be considering the Himalayan Climbing Expedition datasets available on GitHub.
[Himalayan Expedition Datasets](https://github.com/rfordatascience/tidytuesday/tree/main/data/2020/2020-09-22)

# Create the connection
```{r}

con_exp <- DBI::dbConnect(duckdb::duckdb(), dbdir = "exp_db")

```

# Read in the raw data from Github
```{r}

members <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-09-22/members.csv')
expeditions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-09-22/expeditions.csv')
peaks <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-09-22/peaks.csv')


```

```{r}
glimpse(peaks)
```

```{r}
summary(peaks)

# Nothing seems like it needs to be removed.
```

```{r}
glimpse(expeditions)
```

```{r}
glimpse(expeditions)
```

```{r}
summary(expeditions)

#Don't see any issue in keeping everything for now, besides peak_name 
# which will be redundant after join.
```

```{r}
glimpse(members)
```

```{r}
summary(members)

# Remove some heavily NA columns, and columns that can be gotten from a join 
# (year, etc)
```

```{r}
# Remove some uninteresting/heavily null columns in members.
# Remove also columns that we will be able to pull from the peaks or expeditions
# table with a join to reduce redundancy when joining (peak_name, peak_id, year,
# and season can all be obtained from joining with expeditions and peak)

# Also remove duplicate key which won't play nice with SQL

members_clean <- members %>% select(!c('highpoint_metres','death_cause','death_height_metres','injury_type','injury_height_metres','peak_name','peak_id','year','season')) %>% 
  filter(member_id!='KANG10101-01')
  

glimpse(members_clean) #looks good
```

```{r}
# Must remove related datapoints from expeditions that we did in above chunk
# for members. This is some weird duplicate point that messes things up.
# also remove "peak_name" column. This can be obtained by joining with peaks
# table, and will be redundant when we do that join

expeditions_clean <- expeditions %>% 
  select(-peak_name) %>% 
  filter(expedition_id != 'KANG10101')
```

```{r}
# Write CSV's to read from for database

write_csv(peaks,'peaks.csv')
write_csv(expeditions_clean, 'expeditions.csv')
write_csv(members_clean,'members.csv')
```

### ---------------------------------------------------------------------------

# Designing the database

```{sql connection=con_exp}

DROP TABLE IF EXISTS members;
DROP TABLE IF EXISTS expeditions;
DROP TABLE IF EXISTS peaks;
```


# Create a table from the members dataset
```{sql connection= con_exp}

CREATE TABLE members (
  expedition_id VARCHAR(20) NOT NULL,
  member_id VARCHAR(30) NOT NULL,
  sex CHAR(1) DEFAULT NULL,
  age SMALLINT DEFAULT NULL,
  citizenship VARCHAR(100) DEFAULT NULL,
  expedition_role VARCHAR(100) DEFAULT NULL,
  hired BOOLEAN DEFAULT NULL,
  success BOOLEAN DEFAULT NULL,
  solo BOOLEAN DEFAULT NULL,
  oxygen_used BOOLEAN DEFAULT NULL,
  died BOOLEAN DEFAULT NULL,
  injured BOOLEAN DEFAULT NULL,

  PRIMARY KEY (member_id)
);

```

# Create a table from the expeditions dataset
```{sql connection= con_exp}

CREATE TABLE expeditions (
  expedition_id VARCHAR(20) NOT NULL,
  peak_id VARCHAR(10) NOT NULL,
  year SMALLINT DEFAULT NULL,
  season VARCHAR(20) DEFAULT NULL,

  basecamp_date DATE DEFAULT NULL,
  highpoint_date DATE DEFAULT NULL,
  termination_date DATE DEFAULT NULL,

  termination_reason VARCHAR(255) DEFAULT NULL,
  highpoint_metres INT DEFAULT NULL,

  members INT DEFAULT NULL,
  member_deaths INT DEFAULT NULL,
  hired_staff INT DEFAULT NULL,
  hired_staff_deaths INT DEFAULT NULL,

  oxygen_used BOOLEAN DEFAULT NULL,
  trekking_agency VARCHAR(100) DEFAULT NULL,

  PRIMARY KEY (expedition_id)
);

```

# Create a table from the peaks dataset
```{sql connection= con_exp}

CREATE TABLE peaks (
  peak_id VARCHAR(10) NOT NULL,
  peak_name VARCHAR(100) DEFAULT NULL,
  peak_alternative_name VARCHAR(150) DEFAULT NULL,
  height_metres INT DEFAULT NULL,
  climbing_status VARCHAR(50) DEFAULT NULL,
  first_ascent_year SMALLINT DEFAULT NULL,
  first_ascent_country VARCHAR(150) DEFAULT NULL,
  first_ascent_expedition_id VARCHAR(20) DEFAULT NULL,

  PRIMARY KEY (peak_id)
);

```

# Copy each table into the database
```{sql connection= con_exp}

COPY members FROM 'members.csv' (FORMAT CSV, HEADER, NULL 'NA');

```

```{sql connection=con_exp}

COPY expeditions FROM 'expeditions.csv' (FORMAT CSV, HEADER, NULL 'NA');

```

```{sql connection=con_exp}

COPY peaks FROM 'peaks.csv' (FORMAT CSV, HEADER, NULL 'NA');

```

# Check database
```{sql connection=con_exp}
SHOW DATABASES;
```

# Check table variables
```{sql connection=con_exp}
DESCRIBE peaks;
```

```{sql connection=con_exp}
DESCRIBE expeditions;
```

```{sql connection=con_exp}
DESCRIBE members;
```

### ---------------------------------------------------------------------------

# Querying data using SQL

# Laura

Average Peak Height by Climbing Status: Compares the average heigh of peaks across different climbing status categories. 
```{sql connection=con_exp}
SELECT climbing_status, AVG(height_metres) AS avg_height
FROM peaks
GROUP BY climbing_status
ORDER BY avg_height DESC;
```

We can see that helps talled peaks are actually more likely to be climbed. This suggests that elevation may not be a key factor in difficulty. Some peaks may remain unclimbed due to factors other than height, such as accessibility or restrictions.


Success Over Time: Examines how the probability of a successful climb changes across years.

```{sql connection=con_exp}

SELECT 
  e.year,
  COUNT(*) AS total_members,
  SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) AS successful_members,
  ROUND(
    SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) * 1.0 / COUNT(*), 
    3
  ) AS success_rate
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
GROUP BY e.year
ORDER BY e.year;
```

We can see that the success rate has a generally upward trend over time with a larger portion of climbers summiting each year. This suggests that this may be due to improvements in gear, technology or training. 


Success VS Oxygen Used: Compares the success rates of those who used supplemental oxygen on their climb vs those who did not. 

```{sql connection=con_exp}

SELECT 
  m.oxygen_used,
  COUNT(*) AS total_climbers,
  SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) AS successes,
  ROUND(
    SUM(CASE WHEN m.success = TRUE THEN 1 ELSE 0 END) * 1.0 / COUNT(*), 
    3
  ) AS success_rate
FROM members m
GROUP BY m.oxygen_used
ORDER BY m.oxygen_used;

```

This output tells us that climbers who used supplemental oxygen (oxygen_used = TRUE) had a much higher success rate that those who didn't. This also tells us that many more climbs have been done without the use of oxygen as the number of total climbers that didn't use oxygen is substantially higher. This could be due to limits in oxygen accessibility over time. 


JOIN - Top Countries Participating: Displays which countries are most involved in Himalayan expeditions. 

```{sql connection=con_exp}

SELECT 
  m.citizenship,
  COUNT(*) AS num_climbers,
  COUNT(DISTINCT m.expedition_id) AS expeditions_participated
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
GROUP BY m.citizenship
ORDER BY num_climbers DESC
LIMIT 10;

```
This analysis shows how participation in Himalayan expeditions varies by country, highlighting which nations contribute the most climbers and how widely they are involved across expeditions. Nepal appears as the largest participator, which makes sense given the location of the Himalayas. Also, climbers from Nepal are also often guides which will also contribute to this number. 

#Tristan

Get an idea of the spread of hired staff vs recreational climbers in the expeditions

```{sql connection=con_exp}
/*Gets max, min, and mean number of members (recreational climbers) and staff
across all expeditions*/

SELECT 
  MAX(members) AS max_members, MIN(members) AS min_members, AVG(members) AS average_members,
  MAX(hired_staff) AS max_staff, MIN(hired_staff) AS min_staff, AVG(hired_staff) AS average_staff
  FROM expeditions

```

We see a pretty big spread, so we'll probably have to dive deeper.  

Check what peaks have the lowest average member to staff ratio for expeditions. We will have to exclude expeditions with 0 members or 0 staff to get
a meaningful result. We will check a count how many of these there are after to get
an idea of how much of the data we're excluding

```{sql connection=con_exp}

SELECT p.peak_name, AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
FROM expeditions e
JOIN peaks p
  ON e.peak_id = p.peak_id
WHERE members != 0 AND hired_staff != 0
GROUP BY p.peak_name
ORDER BY member_staff_ratio 

```
We have around 10,000 rows total, so the above query only acts on about 63% of our data. I still think this is enough to produce some interesting results
```{sql connection=con_exp}
/* Check how many expeditions had no staff or no recreational members. This is the number
of data points excluded in our above query*/

SELECT count(*) as count
FROM expeditions 
WHERE members = 0 OR hired_staff = 0

```
Let's look a bit closer now at success rates. We will exclude the same rows we excluded above to stay consistent

```{sql connection=con_exp}
  SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name
  ORDER BY success_rate DESC
```
Seems like a lot of them are 1 or 0. This could mean there are only a few (or one) expeditions.
We will still use this for plotting, as it might be interesting to see how the member/staff ratio looks for some of these. But for interest sake lets filter these 1's and 0's out to see the variation besides these cases.

```{sql connection=con_exp}
/*Use above query as a subquery, then select all where the success rate is not 0 or 1*/

SELECT *
FROM (
  SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name
) t
WHERE success_rate NOT IN (0, 1)
ORDER BY success_rate;


```



### ---------------------------------------------------------------------------

# Creating data visualizations

#Laura

Death Rate of Himalayan Climbers Over Time: Shows how the death rate of Himalayan climbers has changed over the years. 

```{sql connection=con_exp}

/*Query to obtain the death rate of number of climbers for each year*/

SELECT 
  e.year,
  AVG(CASE WHEN m.died = TRUE THEN 1 ELSE 0 END) AS death_rate,
  COUNT(*) AS total_climbers
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
WHERE e.year IS NOT NULL
GROUP BY e.year
HAVING COUNT(*) > 50
ORDER BY e.year;
```

```{r}
# entering previous sql code into r in order to run data visualization with 
# ggplot

death_trend_data <- dbGetQuery(con_exp, "
SELECT 
  e.year,
  AVG(CASE WHEN m.died = TRUE THEN 1 ELSE 0 END) AS death_rate,
  COUNT(*) AS total_climbers
FROM members m
JOIN expeditions e
  ON m.expedition_id = e.expedition_id
WHERE e.year IS NOT NULL
GROUP BY e.year
HAVING COUNT(*) > 50
ORDER BY e.year
")

```

```{r}
# plotting data using ggplot with minimal aesthetics for increased simplicity 
# and understanding
ggplot(death_trend_data, aes(x = year, y = death_rate)) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Death Rate of Himalayan Climbers Over Time",
    x = "Year",
    y = "Death Rate"
  ) +
  theme_minimal()

```
By plotting the proportion of climbers who died each year, we can see that the death rate has steadily declined since 1970, which makes sense as there has been upgrades to technology as well as more guides present on these expeditions. The data increases steadily at the start which makes sense given the smaller number of climbers in the beginning years as well as the lack of knowledge of the climb. 


#Tristan

#Success Rate vs Member to Staff ratio

My interest here is to see if there is any negative correlation between these two variables. 
I expect that there will be some, but not extremely strong, as what really matters
is the difficulty of the peak, the experience of the people, etc. But presumably more professional staff members present should result in more success. So I'm imagining that 
a smaller member to staff ratio (meaning less members per staff) should result in more success

```{sql connection=con_exp}
/*Query to get both the success rate and the member-staff ratio for all peaks 
without 100% or 0% success rate*/

 SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate,
    AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name

```

```{r}
#Convert the result of the above query into a data.frame

success_rate_member_ratio_data <- dbGetQuery(con_exp, "
SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate,
    AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name")

```

```{r}
#Plot our variables of interest from our convereted data.frame

ggplot(success_rate_member_ratio_data, aes(x = member_staff_ratio, y = success_rate)) +
  geom_point() +
  geom_smooth(se = FALSE)+
labs(
    title = "Success Rate of Expeditions vs Recreational Climber/Staff Ratio",
    x = "Recreational Climber/Staff Ratio",
    y = "Success Rate"
  ) +
  theme_minimal()
```

Check out filtered version too

```{r}

success_rate_member_ratio_data_filtered <- dbGetQuery(con_exp, "
 SELECT * FROM
 (SELECT
    p.peak_name,
    SUM(CASE WHEN e.termination_reason LIKE '%Success%' THEN 1 ELSE 0 END) * 1.0 /
    COUNT(*) AS success_rate,
    AVG(e.members * 1.0 / e.hired_staff) AS member_staff_ratio
  FROM expeditions e
  JOIN peaks p
    ON e.peak_id = p.peak_id
  WHERE (e.members != 0 AND e.hired_staff != 0)
  GROUP BY p.peak_name
 ) t
WHERE success_rate NOT IN (0, 1)
 ")

```

```{r}
ggplot(success_rate_member_ratio_data_filtered, aes(x = member_staff_ratio, y = success_rate)) +
  geom_point() +
  geom_smooth(se = FALSE)+
labs(
    title = "Success Rate of Expeditions vs Recreational Climber/Staff Ratio",
    x = "Recreational Climber/Staff Ratio",
    y = "Success Rate"
  ) +
  theme_minimal()
```

There isn't too much correlation evident in these plots. If there is, it seems to actually be slightly positive. We do see many more 100% succesful expeditions with higher member to staff ratios, so these expeditions were succesful even with less staff. Overall it seems that the most important characteristics are likely things like the peak difficulty, the experience of the climbers, etc. This calculation I have done here is far too broad it seems to capture anything meaningful. I would like maybe to dive deeper on my own and see if I can discover some more about this dataset.
### ---------------------------------------------------------------------------

```{r}
# terminate the SQL connection

DBI::dbDisconnect(con_exp, shutdown = TRUE)

```

# Sources

-   web link/document title/LLM prompt

-   web link/document title/LLM prompt

-   web link/document title/LLM prompt

-   add more if needed

### ---------------------------------------------------------------------------
